Report on laboratory assignments 1 and 2 for EDA282, \\ Parallel Computer Organization and Design

### 2014-10-14 lab/report/include/introduction.tex
written some text
new figures
wrote and added appendix
correcting

\section{Introduction}
\label{sec:int}
These laboratory assignments aim to increase our knowledge of different cache coherence protocols. The goal is to simulate cache coherence protocols (CCP) on a modeled cache coherent multi-core system using diverse benchmarks. Three cache coherence protocols are targeted; MSI, MESI, and MESI-MG, where the last protocol is an modified version of MESI optimized for migratory sharing. The system model have a shared Last Level Cache (LLC), and each processor have a private L1 cache. Snooping based write-invalidate cache coherence protocols are used by the private L1 caches to ensure coherence, and a simple bus stand for the communication between LLC and private caches. The system is modeled by the Pin tool MultiCacheSim and Pin. In listing~\ref{lst:setup} is the command used for simulating a CCP on the modeled system with a chosen benchmark.

\begin{lstlisting}[label=lst:setup, basicstyle=\scriptsize, caption={Command used for simulation.}]
#> ./pin -mt -t ./source/tools/MultiCacheSim/obj-intel64/mcs.so -csize 65536 -bsize 64 -assoc 8 -llc -llcsize 4194304 -llcbsize 64 -llcassoc 16 -numcaches 8 -protos ./source/tools/MultiCacheSim/obj-intel64/mcs_<CCP>.so -- <benchmark> [#threads,..]
\end{lstlisting}

>>>>

\section{Introduction}
\label{sec:int}
These laboratory assignments aim to increase our knowledge of different cache coherence protocols. To reach this goal, we simulate cache coherence protocols (CCP) on a modeled cache coherent multi-core system using diverse benchmarks. Three cache coherence protocols are targeted; MSI, MESI, and MESI-MG, where the last protocol is an modified version of MESI optimized for migratory sharing. The system model have a shared Last Level Cache (LLC), and each processor have a private L1 cache. Snooping based write-invalidate cache coherence protocols are used by the private L1 caches to ensure coherence, and a simple bus stand for the communication between LLC and private caches. The system is modeled by the Pin tool MultiCacheSim and Pin. In listing~\ref{lst:setup} is the command used for simulating a CCP on the modeled system with a chosen benchmark. To relate the simulation results to excecution time, Eq. \ref{eq:cycles} is used.

\begin{lstlisting}[label=lst:setup, basicstyle=\scriptsize, caption={Command used for simulation.}]
#> ./pin -mt -t ./source/tools/MultiCacheSim/obj-intel64/mcs.so -csize 65536 -bsize 64 -assoc 8 -llc -llcsize 4194304 -llcbsize 64 -llcassoc 16 -numcaches 8 -protos ./source/tools/MultiCacheSim/obj-intel64/mcs_<CCP>.so -- <benchmark> [#threads,..]
\end{lstlisting}
\begin{equation}
	$$\label{eq:cycles}$$
	Cycles = 1 * (WriteHit+ReadHit) + 30 * (readMissesServicedByModified + invalidationsSent)$$ $$+ 20 * LLCHit + 50 * LLCMiss
\end{equation}

### 2014-10-14 lab/report/include/lab2.tex
correcting

\section{Lab 2}
\label{sec:lab2}
For the second part of the lab the task is to parallelize a serial implementation of the Red-Back Gauss-Seidal algorithm. This is done using OpenMP (Open \textbf Multi-\textbf Processing) which is an API for forming multi-threaded programs in C, C++, and Fortran. This parallel version is evaluated using the same simulator and CCP:s as in \nameref{sec:lab1}.

\subsection{Task 1}
The program reads a matrix on the form $n \times n$ containing data points from a file. To make the program simpler and avoid special cases a padding is added to the matrix changing its dimensions to $(n+2) \times (n+2)$. Then every other data point, here by called red dots, is looped through. For every red dot an average of itself and its neighbors to the west, south, north, and east is calculated and written as the dots new value. For all the other data points, called black dots, the same thing is performed. The difference of the dots new and old value is added together and the process is repeated until this accumulated difference of averages, divided by $n^2$, is smaller than $0.01$. After the program finishes its execution the matrix have converged to an average of all red and black dots.

The distribution of the red and black dots forms a chess pattern. This means that when we calculate the average for a dot and its neighbors, all the neighbors are of the other color. As such, when going through for example the red dots there is no need to invalidate any other dots because the black dots are only read and you are the only one accessing your red dot.

\subsection{Task 2}
\label{subsec:lab2:task2}
In listing~\ref{lst:solve} a parallelized version of the function \texttt{void Solve(double **A)} are shown. The parallelization have been done using directives from the OpenMP API. Only this function is shown as the other parts of the program for reading the input file and printing to the console is not included in the performance analysis. The approach taken for making the function execute in parallel is to distribute the lines of the matrix evenly between all threads. This is done twice, once for the red dots and once for the black. Only one thread resets \texttt{diff} to zero on line 5 and test whether the program is finished or not on line 25. At each for loop that goes through the red and black dots the directive \texttt{reduction(+:diff)} is used which causes each thread to have its own private version of \texttt{diff} that all gets added together after the for loop.

>>>>

\section{Lab 2}
\label{sec:lab2}
For the second part of the lab the task is to parallelize a serial implementation of the Red-Back Gauss-Seidal algorithm. This is done using OpenMP (Open \textbf Multi-\textbf Processing) which is an API for forming multi-threaded programs in C, C++, and Fortran. This parallel version is evaluated using the same simulator and CCPs as in \nameref{sec:lab1}.

\subsection{Task 1}
The program first reads a matrix on the form $n \times n$ containing data points from a file. To make the program simpler and avoid special cases padding is added changing it's dimensions to $(n+2) \times (n+2)$. Then every other data point, here by called red dots, is looped through. For every red dot, an average of itself and it's neighbors to the west, south, north, and east is calculated and written as the dot's new value. The same procedure is performed on all the other data points, called black dots. The difference of the dots new and old values is added together and the process of calculating new averages is repeated until this accumulated difference of averages, divided by $n^2$, is smaller than $0.01$. When the program finishes the matrix have converged to an average of all red and black dots.

The distribution of the red and black dots forms a chess pattern. This means that when the average for a dot and it's neighbors is calculated, all the neighbors are of the other color. As such, when calculating the average for example the red dots there is no need to invalidate any other dots (than red dots???) because the black dots are only read and you are the only one accessing your red dot.

\subsection{Task 2}
\label{subsec:lab2:task2}
In listing~\ref{lst:solve} a parallelized version of the function \texttt{void Solve(double **A)} is shown. The approach taken to parallilize the function is to distribute the rows of the matrix evenly between all threads. This is done twice, once for the red dots and once for the black. Only one thread resets \texttt{diff} to zero on line 5 and test whether the program is finished or not on line 25. At each for loop that goes through the red and black dots the directive \texttt{reduction(+:diff)} is used which causes each thread to have its own private version of \texttt{diff} that all gets added together after the for loop.

### 2014-10-14 lab/report/main.tex
written some text
write
added stupid table. crapy shiet
task4
correcting

\usepackage[margin=2cm]{geometry}
\include{packages}
\usepackage[toc,page]{appendix}
\usepackage{color}
%MACROS---
\newcommand{\reffig}[1]{Fig. \ref{#1}}
\newcommand\xput[2][0.5]{%
    \rule{#1\linewidth}{0pt}\makebox[0pt][c]{#2}\hfill}

\title{Report on Lab 1 \& Lab 2 in EDA282, Parallel Computer Organization and Design}
\date{October 2014}
\author{Dan Larsson, Jonas Hemlin
\\Master Program of Embedded Electronic Systems Design\\
Chalmers University of Technology, Gothenburg, Sweden\\
Email: \{larsdan, jhemlin\}@student.chalmers.se}

@@

\begin{document}
\maketitle
$$ Cost = 1 * (WriteHit+ReadHit) + 30 * (readMissesServicedByShared + invalidationsSent)$$ $$+ 20 * LLCHit + 50 * LLCMiss $$
\input{include/introduction}
\input{include/lab1}
\input{include/lab2}
\input{include/discussion}
\newpage
\begin{appendices}
\section{Parallelized program for Lab 2}
\centerline{
\begin{minipage}{1.0\textwidth}
\lstinputlisting[caption={\textit{Parallel version of original} \texttt{void Solve(double **A).}}, label={lst:solve}]{include/solve.c}
\end{minipage}
}
\end{appendices}
\end{document}

>>>>

\usepackage[margin=2cm]{geometry}
\include{packages}
\usepackage[toc,page]{appendix}
\usepackage{color}
\usepackage{lscape}
\usepackage{chngpage}
\usepackage{float}
%MACROS---
\newcommand{\reffig}[1]{Fig. \ref{#1}}
\newcommand\xput[2][0.5]{%
    \rule{#1\linewidth}{0pt}\makebox[0pt][c]{#2}\hfill}

\title{Report on laboratory assignments 1 and 2 for EDA282, \\ Parallel Computer Organization and Design}
\date{October 2014}
\author{Dan Larsson, Jonas Hemlin
\\Master Program of Embedded Electronic Systems Design\\
Chalmers University of Technology, Gothenburg, Sweden\\
Email: \{larsdan, jhemlin\}@student.chalmers.se}

@@

\begin{document}
\maketitle
\input{include/introduction}
\input{include/lab1}
\input{include/lab2}
\input{include/discussion}
\newpage
\begin{appendices}
\input{include/lab2appendice}
\end{appendices}
\end{document}
